---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = " "
)
library(googlesheets4)
library(magrittr)
x <- gs4_auth(email = readLines("~/mail"))
wk <- read_sheet(readLines("~/sheet"))
```

# ur-first-5k

Running your first 5K by closing GitHub issues.

## About

This repo is here to help you get started with running by training you for your first 5k 🎉

### How it works

Each week during 10 weeks, you'll be assigned 3 issues, each with one running session, plus an optional "cross training session" (bike/yoga/pilate/stretching...).

> Note that o can assign you an issue only once you've accepted being a collaborator to the repo. (check your email for the link).
Each issue is on the form of "YourGithubName - CODE-Type", where code is Week number (1:10) + Day Number (1:7), and Type is a short description of the workout. 
When you've completed the workout, you can come and close the issue. 
Feel free to add a comment on the issue or a question if you want to exchange about your experience.

You can also join a general discussion at [ur-first-5k/issues/1](https://github.com/ColinFay/ur-first-5k/issues/1)

Just so that you only focus on the getting the current week done, you'll only get the trainig plan one week at a time :) 

Weeks are organized around 5 types of sessions: 

+ "Easy run", where you don't really focus on your pace or on running the whole time. 
Feel free to mix this run with walking and run at an easy pace.
This run is more about enjoying the run and having a good time. 

+ "Interval run", where you will do intervals of running and walking. 
This session will help you build a better strength and cardio.

+ "Steady Run" is a form of intervals that mixes easy pace & a more fast pace, running the whole time.
Imagine it as a interval session on steroids (instead of walking/running, you're doing easy-run/fast-run).

+ "Long run", with a slower pace than the interval runs, but with a longer time spent running than easy runs. 

+ "Recovery run" often comes an hard session. 
The idea with this session is to keep the machine pumping without tiring your body too much.

+ A "cross training" session: feel free to add to the mix a day of any other sport (bike) or doing aerobic/yoga/pilate. 

### Organizing your week

Given that you are starting to run, you'll need one or two days to rest between each session.
I suggest doing your session on Tuesday, Thursday and Sunday, but feel free to switch to another organisation that will fit your personal preferences.

### Some advices

+ When we start running, we usually go too fast and push too hard as soon as we start running. 
This changes when we're getting to know ourselves and control our good pace, so the first advice will be: don't try to push too hard when you will start :)

+ When running at an "easy pace", try to keep a pace where you could still be talking. 
The goal for this plan is to give you enough ease so that you can run 5k in the end. 
In other words, we don't really care about speed--try to keep an easy pace so that you don't become breathless after 30 seconds of running.

+ During the intervals, the "recovery walk" should still be an active walk: the idea is to keep your cardio up while allowing you to get your breath back. 
In other words, when doing the intervals, don't stop being active between the running segments.

## Plan

### Week one

```{r echo = FALSE}
one <- wk[wk$week == 1, 2:ncol(wk) ]
for (i in 1:nrow(one)){
  loc <- one[i, ]
  cli::cat_rule(loc$day)
  cli::cat_bullet(loc$code, bullet = "arrow_right")
  cli::cat_line(loc$workout)
  cli::cat_line()
}
```

### Week two

```{r echo = FALSE}
one <- wk[wk$week == 2, 2:ncol(wk) ]
for (i in 1:nrow(one)){
  loc <- one[i, ]
  cli::cat_rule(loc$day)
  cli::cat_bullet(loc$code, bullet = "arrow_right")
  cli::cat_line(loc$workout)
  cli::cat_line()
}
```


## External links

[The RStats Strava Club](https://www.strava.com/clubs/rstats)

[10-Minute Guided Stretching Routine for Runners](https://www.youtube.com/watch?v=W9YtvgljfyY)

[How to stretch after a run](https://www.nhs.uk/live-well/exercise/how-to-stretch-after-a-run/)

[8 Essential Yoga Poses For Runners](https://www.runnersworld.co.za/training/8-yoga-poses-for-runners/)

[7 Pilates Exercises That Build a Stronger Core](https://www.runnersworld.com/training/a20865857/pilates-exercises-for-runners/)